Iterative Reduction: Parsing and Interpreting Code Without Trees

### 2024-12-16 Paper/paper.tex
comments on the third example

The initial $\beta$-reducer is a function
\[ \hb \colon \Sigma\times\Sigma \partial \Sigma\times(\bbZ\times\bbZ\to\bbZ)\times(\values\times\values\partial\values) \]
Our alphabet of types is $\Sigma=\set{\NUM,\OP,\LPAREN,\RPAREN,\OP\NUM,\RPAREN\NUM}$, define $\oP=\oS\times\values$ to be the set of all possible terms (a term consists of a type, a priority, and a value).
Since terms now consist of three parts, we write them as $\sigma_p(v)$, where $\sigma$ is the type; $p$ is the priority; and $v$ the value.

@@

Our initial $\beta$-reducer will is defined as follows:
\begin{itemize}
    \item $\hb(\NUM,\OP)=(\OP\NUM,\snd,(v,f\mapsto(v,f))$
    \item $\hb(\OP\NUM,\NUM)=(\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\OP\NUM,\OP\NUM)=(\OP\NUM,\snd,((v,f),(u,g)\mapsto (f(v,u),g))$
    \item $\hb(\NUM,\RPAREN)=(\RPAREN\NUM,\snd,(v,\_\mapsto v))$
    \item $\hb(\OP\NUM,\RPAREN\NUM)=(\RPAREN\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\LPAREN,\RPAREN\NUM)=(\NUM,\fst,(\_,v\mapsto v))$
\end{itemize}

So taking our previous example of \ttt{2*(1+3)+4}: we first must convert it into tokens acceptable for $\beta$-reduction (via lexing).
So it is converted to
\[ \NUM_\infty(2)\ \OP_2(*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \]
The derivation is then
\begin{gather*}
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \OP\NUM_1(1,+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \OP\NUM_1(1,+)\ \RPAREN\NUM_0(3)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \RPAREN\NUM_0(4)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \NUM_\infty(4)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \OP\NUM_1(4,+)\ \NUM_\infty(4) \\
    \OP\NUM_2(8,+)\ \NUM_\infty(4) \\
    \OP\NUM_2(8,+)\ \NUM_0(4) \\
    \NUM_0(12)
\end{gather*}

But this of course is not capable of handling the semantics necessary for programming languages.
We need a notion of a state, and a somewhat more refined algorithm.

>>>>

\yoni{I would consider using the word "kinds" instead of "types". "kinds" is not great either, but types are stuff like int, real, int->int, which i am not sure is appropriate here.}

The initial $\beta$-reducer is a function
\[ \hb \colon \Sigma\times\Sigma \partial \Sigma\times(\bbZ\times\bbZ\to\bbZ)\times(\values\times\values\partial\values) \]
Define $\oP=\oS\times\values$ to be the set of all possible terms,
that is, 
a term consists of a type, a priority, and a value.
Since terms now consist of three parts, we write them as $\sigma_p(v)$, where $\sigma$ is the type; $p$ is the priority; and $v$ the value.

@@

\begin{example}
Define $\Sigma=\set{\NUM,\OP,\LPAREN,\RPAREN,\OP\NUM,\RPAREN\NUM}$.
Our initial $\beta$-reducer will is defined as follows:
\begin{itemize}
    \item $\hb(\NUM,\OP)=(\OP\NUM,\snd,(v,f\mapsto(v,f))$
    \item $\hb(\OP\NUM,\NUM)=(\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\OP\NUM,\OP\NUM)=(\OP\NUM,\snd,((v,f),(u,g)\mapsto (f(v,u),g))$
    \item $\hb(\NUM,\RPAREN)=(\RPAREN\NUM,\snd,(v,\_\mapsto v))$
    \item $\hb(\OP\NUM,\RPAREN\NUM)=(\RPAREN\NUM,\snd,((v,f),u\mapsto f(v,u)))$
    \item $\hb(\LPAREN,\RPAREN\NUM)=(\NUM,\fst,(\_,v\mapsto v))$
\end{itemize}

\yoni{I don't really understand it. For example, what is 
$(v,f\mapsto(v,f))$? Perhaps introducing some more definitions
and details can help clarify?
}
\yoni{What happened to the squares around the tokens in this example?}

So taking our previous example of \ttt{2*(1+3)+4}: we first must convert it into tokens acceptable for $\beta$-reduction (via lexing).
So it is converted to
\[ \NUM_\infty(2)\ \OP_2(*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \]

\yoni{Here we are missing a formal assignment of priorities.}

The derivation is then
\begin{gather*}
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \NUM_\infty(1)\ \OP_1(+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \OP\NUM_1(1,+)\ \NUM_\infty(3)\ \RPAREN_0\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \OP\NUM_1(1,+)\ \RPAREN\NUM_0(3)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \LPAREN_\infty\ \RPAREN\NUM_0(4)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \NUM_\infty(4)\ \OP_1(+)\ \NUM_\infty(4) \\
    \OP\NUM_2(2,*)\ \OP\NUM_1(4,+)\ \NUM_\infty(4) \\
    \OP\NUM_2(8,+)\ \NUM_\infty(4) \\
    \OP\NUM_2(8,+)\ \NUM_0(4) \\
    \NUM_0(12)
\end{gather*}

\end{example}
But this of course is not capable of handling the semantics necessary for programming languages.
We need a notion of a state, and a somewhat more refined algorithm.
